Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
more code comment

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for 4 non zeroes of
the matrix at once. The 4 floats coming from each matrix four 4
consecutive non zeroes are naturally grouped within each lane of the
SIMD vector. One can use a swizzle operation to broadcast the 4
elements of a single matrix to take the whole vector. (The values
$A_h^1,A_h^2,A_h^3,A_h^4$ are broadcasted in Figure~\ref{code:mat_mul}).

Loading the entries from the {\tt x} vector is a little bit more
complicated. First the columns pointers of 4 different non zeroes are
loaded in a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to unpack for the
  LSB and MSB. Both are mandatory even if one is known not to be
  useful according the documentation of the hardware}. (The column
pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
replicated so that each fill a vector lane using a swizzle
operation. Then each value is multiplied by 4 (because there are 4
vectors) and the offsets $(3,2,1,0)$ is added each lane to obtain the
correct index of the elements of the {\tt x} vector within each lane
using a Fused Multiply-Add operation. Then a {\tt gather} operation is
performed bring the 16 entries of the {\tt x} vector into the SIMD
register. At this point we have a SIMD register where each lane is the
four vector entries for a non zero element. Using a broadcast
permutation, one can replicate an entry of the four vectors accross
each lane.

>>>>

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for 4 non zeroes of
the matrix at once. The 4 floats coming from each matrix four 4
consecutive non zeroes are naturally grouped within each lane of the
SIMD vector. One can use a swizzle operation to broadcast the 4
elements of a single matrix to take the whole vector. (The broadcast of 
$A_h^1,A_h^2,A_h^3,A_h^4$ is shown in Figure~\ref{code:mat_mul}.)

Loading the entries from the {\tt x} vector is a little bit more
complicated. First the columns pointers of 4 different non zeroes are
loaded in a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to unpack for the
  LSB and MSB. Both are mandatory even if one is known not to be
  useful according the documentation of the hardware}. (The column
pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
replicated so that each fill a vector lane using a swizzle
operation. Then each value is multiplied by 4 (because there are 4
vectors) and the offsets $(3,2,1,0)$ is added each lane to obtain the
correct index of the elements of the {\tt x} vector within each lane
using a Fused Multiply-Add operation. Then a {\tt gather} operation is
performed bring the 16 entries of the {\tt x} vector into the SIMD
register. At this point we have a SIMD register where each lane is the
four vector entries for a non zero element. Using a broadcast
permutation, one can replicate an entry of the four vectors accross
each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
Figure~\ref{code:mat_mul}.)

Now that $A_h^1,A_h^2,A_h^3,A_h^4$ and $x_m^1,x_m^2,x_m^3,x_m^4$ are
properly laid out in the vectors, a partial value for $y_i$ can be
computed. Once all the non zeroes of the row have been processed, the
exact $y_i$ value can be sent to memory using a the No Read No Global
Order hint to optimize traffic on the memory bus.
